%% -*- coding:utf-8 -*-
The 13th International Conference on Head-Driven Phrase Structure Grammar (2006) was held in Varna
and organized by the Linguistic Modelling Laboratory of the Institute for Parallel Processing of the
Bulgarian Academy of Sciences in Sofia.

The conference featured 2 invited talks, 16 papers, and 4 posters
selected by the program committee 
(Anne Abeillé,
Raul Aranovich,
Emily Bender,
Gosse Bouma,
António Branco,
Chan Chung,
Ann Copestake,
Berthold Crysmann,
Elisabeth Engdahl,
Anna Feldman,
Dan Flickinger,
Howard Gregory,
Daniele Godard,
Erhard Hinrichs (chair),
Jong-Bok Kim,
Valia Kordoni,
Ania Kupsc,
Shalom Lappin,
Robert Levine,
Stefan Müller,
Tsuneko Nakazawa,
Petya Osenova,
Gerald Penn,
Luisa Sadler,
Ivan Sag,
Manfred Sailer,
Gautam Sengupta,
Jan-Philipp Soehn (chair),
Jesse Tseng,
Nathan Vaillette,
Stephen Wechsler,
Eun-Jung Yoo,
Larisa Zlatic).

A workshop about \emph{Regularity and Irregularity in Grammar and Language}
was attached to the conference. It featured one invited talk
and 5 papers, selected by the program committee.

In total there were 39 submissions to the main conference and  submissions to the
workshop. %<!--, which gives an acceptance rate of %.-->
We want to thank the respective program committee for putting this nice program together.




Thanks go to Kiril Simov and Petya Osenova, who were in charge of local arrangements.


As in the past years the contributions to the conference proceedings are based on the five page abstract
that was reviewed by the respective program committees, but there is no additional reviewing of the
longer contribution to the proceedings.
To ensure easy access and fast publication we have chosen an electronic format.


The proceedings include all the papers except those by Stefan Müller, Shravan Vasishth, and Frank
van Eynde.
