The 12th International Conference on Head-Driven Phrase Structure Grammar (2005) was held at
the Department of Informatics, University of Lisbon in Portugal.

The conference featured 2 invited talks, 18 papers, 2 alternate papers, and 6 posters
selected by the program committee 
(Raul Aranovich,
Doug Arnold,
Emily Bender,
Olivier Bonami,
António Branco,
Berthold Crysmann,
Anke Holler,
Valia Kordoni,
Palmira Marrafa,
Tsuneko Nakazawa,
Gerald Penn,
Alexander Rosen,
Manfred Sailer (chair),
Gautam Sengupta,
Jesse Tseng,
Stephen Wechsler, and
Shuly Winter). 
A workshop on \emph{Binding Theory and Invariants in Anaphoric Relations} was attached to the conference. It featured one invited talk
and 12 papers, selected by the workshop program committee (Pilar Barbosa,
António Branco (chair),
Rejean Canac-Marquis,
Mary Dalrymple,
Martin Evearert,
Volker Gast,
Lars Hellan,
Ehrard Hinrichs,
Yan Huang,
Tibor Kiss,
Frank Keller,
Valia Kordoni,
Maria Pinango,
Carl Pollard,
Janina Radó,
Eric Reuland,
Jeffrey Runner,
Ivan Sag,
Roland Stuckardt,
Ping Xue).

In total there were 39 submissions to the main conference and 13 submissions to the
workshop. 
We want to thank the respective program committees for putting this nice program together.

Thanks go to António Branco, who was in charge of local arrangements.

As in the past years the contributions to the conference proceedings are based on the five page abstract
that was reviewed by the respective program committees, but there is no additional reviewing of the
longer contribution to the proceedings.
To ensure easy access and fast publication we have chosen an electronic format.

The proceedings include all the papers except those by 
Frank Keller and Theodora Alexopoulou, Stefan Müller and Eric Reuland. Nurit Melnik submitted
an extended abstract, the full paper will appear in a Research on Language and Computation.
