%% -*- coding:utf-8 -*-
The 11th International Conference on Head-Driven Phrase Structure Grammar (2004) was held at
the Center for Computational Linguistics, Katholieke Universiteit Leuven in Belgium.

The conference featured 3 invited talks, 20 papers, and 1 alternate paper,
selected by the program committee (Gosse Bouma, Jonathan Ginzburg, Jong-Bok Kim, Tibor Kiss, Kordula de Kuthy, Bob Levine, Rob Malouf, Philip Miller, Paola Monachesi, Stefan Müller, John Nerbonne, Ivan Sag, Manfred Sailer, Frank Van Eynde, Eun-Jung Yoo, Kei Yoshimoto). 
A workshop on \emph{Semantics in Grammar Engineering} was attached to the conference. It featured one invited talk
and 7 papers, selected by the workshop program committee (Dorothee Beermann, Ash Asudeh, Mary Dalrymple, Markus Egg, Lars Hellan, Jean-Pierre Koenig, Valia Kordoni, Manfred Sailer).

In total there were 32 submissions to the main conference and 7 submissions to the
workshop. 
We want to thank the respective program committees for putting this nice program together.

Thanks go to Frank van Eynde, who was in charge of local arrangements.

As in the past years the contributions to the conference proceedings are based on the five page abstract
that was reviewed by the respective program committees, but there is no additional reviewing of the
longer contribution to the proceedings.
To ensure easy access and fast publication we have chosen an electronic format.

The proceedings include all the papers except those by Farrell Ackerman, Irina Nikolaeva, and Rob Malouf, Ann Copestake, 
Liv Ellingsen, Jonathan Ginzburg, Tibor Kiss, and Adam Przepiórkowski and Alexandr Rosen.
