%% -*- coding:utf-8 -*-
The 25th International Conference on Head-Driven Phrase Structure Grammar (2018) was held at
the University of Tokyo, Komaba Campus.

The conference featured 2 invited talks, 13 papers, and 3 posters selected by the program committee 
(Anne Abeillé,
Doug Arnold,
Daisuke Bekki,
Olivier Bonami,
Francis Bond,
Gosse Bouma,
George Broadwell,
Rui Chaves,
Philippa Cook,
Berthold Crysmann,
Kordula De Kuthy,
Antske Fokkens,
Petter Haugereid,
Fabiola Henri,
Anke Holler,
Gianina Iordachoaia,
Jong-Bok Kim,
Jean-Pierre Koe\-nig,
Yusuke Kubota,
David Lahm,
Robert D. Levine,
Yo Matsumoto,
Nurit Melnik,
Philip Miller,
Stefan Müller,
Tsuneko Nakazawa,
Joanna Nykiel,
Rainer Osswald,
Gerald Penn,
Frank Richter (chair),
Manfred Sailer,
Pollet Samvellian,
Sanghoun Song,
Frank van Eynde,
Stephen Wechsler,
Shûichi Yatabe,
Eun-Jung Yoo).

There was a workshop on \emph{The Clause Structure of Japanese and Korean} with three talks and two
invited talks.

% wie viele?
%In total there were x  submissions to the conference and x submissions to the workshop.
We want to thank the program committees for putting this nice program together.

Thanks go to Shuichi Yatabe, who was in charge of local arrangements, and his assistants Kei Tanigawa,
Tsuneko Nakazawa,
Morine Kondo,
Mayu Kawakita,
Takeshi Kishiyama, and
Fuga Terasaki.
 

As in the past years the contributions to the conference proceedings are based on the five page abstract
that was reviewed by the respective program committees, but there is no additional reviewing of the
longer contribution to the proceedings.
To ensure easy access and fast publication we have chosen an electronic format.

The proceedings include all the papers except the one by Anne Abeillé \& Shrita Hassamal, Gabriel Aguila"=Multner \&
Berthold Crysmann, Makoto Kanazawa, Yusuke Kubota, and Frank Van Eynde. 


