%% -*- coding:utf-8 -*-
The 26th International Conference on Head-Driven Phrase Structure Grammar (2019) was held at
the University of Bucharest.

The conference featured 2 invited talks and 10 papers selected by the program committee 
(Anne Abeillé, 
Doug Arnold, 
Emily Bender, 
Felix Bildhauer, 
Olivier Bonami, 
Francis Bond, 
Gosse Bouma, 
Antonio Branco, 
Rui Chaves, 
Philippa Cook, 
Berthold Crysmann, 
Dan Flickinger, 
Antske Fokkens, 
Petter Haugereid, 
Fabiola Henri, 
Anke Holler, 
Gianina Iordăchioaia, 
Jong-Bok Kim, 
Jean-Pierre Koenig, 
David Lahm, 
Bob Levine, 
Nurit Melnik, 
Philip Miller, 
Stefan Müller, 
Tsuneko Nakazawa, 
Rainer Osswald, 
Petya Osenova (chair), 
Gerald Penn, 
Frank Richter, 
Louisa Sadler, 
Manfred Sailer, 
Pollet Samvellian, 
Jesse Tseng, 
Frank van Eynde, 
Stephen Wechsler, 
Shûichi Yatabe, 
Eun-Jung Yoo).

There was a workshop on \emph{Romance languages} with five talks and one
invited talk.

% wie viele?
%In total there were x  submissions to the conference and x submissions to the workshop.
We want to thank the program committees for putting this nice program together.

Thanks go to Gabriela Bîlbîie and Emil Ionescu, who were in charge of local
arrangements.
 

As in the past years the contributions to the conference proceedings are based on the five page abstract
that was reviewed by the respective program committees, but there is no additional reviewing of the
longer contribution to the proceedings. To ensure easy access and fast publication we have chosen an electronic format.

The proceedings include all the papers of the conference except the ones by 
Anne Abeillé \& Elodie Winckel, 
Gabrielle Aguila-Multner \& Berthold Crysmann,
Antonio Machicao y Priemer \& Paola Fritz-Huechante, 
Nurit Melnik \& Bracha Nir, 
Stefan Müller, Jong-Bok Kim \& Alain Kihm, and Frank Van Eynde, 
who will submit their papers to journals. The
workshop contributions will be published in 2020 in issue 43:1 of Lingvisticæ Investigationes.


