%% -*- coding:utf-8 -*-
The 22th International Conference on Head-Driven Phrase Structure Grammar (2015) was held at the Nanyang Technological University (NTU), Singapore.

The conference featured 3 invited talks, 15 papers, and 4 posters selected by the program committee 
(Anne Abeillé,
    Farrell Ackerman,
    Doug Arnold,
    Emily M. Bender (chair),
    Francis Bond,
    Gosse Bouma,
    George Broadwell,
    Rui Chaves,
    Philippa Cook,
    Ann Copestake,
    Kordula De Kuthy,
    Elisabet Engdahl,
    Dan Flickinger,
    Antske Fokkens,
    Danièle Godard,
    Petter Haugereid,
    Fabiola Henri,
    Anke Holler,
    Jong-Bok Kim,
    Jean-Pierre Koenig,
    Anna Kupsc,
    Bob Levine,
    Janna Lipenkova,
    Rob Malouf,
    Nurit Melnik,
    Philip Miller,
    Tsuneko Nakazawa,
    Joanna Nykiel,
    Gerald Penn,
    Adam Przepiorkowski,
    Frank Richter,
    Louisa Sadler,
    Pollet Samvellian,
    Sanghoun Song,
    Jesse Tseng,
    Steve Wechsler,
    Shûichi Yatabe and
    Eun-Jung Yoo).

A workshop on \emph{Verb Classes and the Scale of Change in Affected Arguments}
was attached to the conference. The workshop had six invited speakers and 
five regular papers. The workshop program was put together by František Kratochvíl and Joanna Ut-Seong Sio.

% wie viele?
%In total there were x  submissions to the conference and x submissions to the workshop.
We want to thank the respective program committees for putting this nice program together.

Thanks go to Francis Bond, who was
in charge of local arrangements, and his assistants Sanghoun Song, Michael Goodman,  Luis Morgado da Costa.
 

As in the past years the contributions to the conference proceedings are based on the five page abstract
that was reviewed by the respective program committees, but there is no additional reviewing of the
longer contribution to the proceedings.
To ensure easy access and fast publication we have chosen an electronic format.

The proceedings include all the papers except the one by Sanghoun Song, Chen Bo, Joanna Sio Ut
Seong, and Francis Bond titled \emph{An HPSG-based Analysis of Resultative Compounds in Chinese},
the one by Frank van Eynde, which is published in his book on predication, and the ones by I Wayan
Arka, Kazuko Yatsuhiro, Juwon Lee, and Hans
Uszkoreit. Most of the workshop contributions will be published in a separate volume and therefore
are also not included.



