%% -*- coding:utf-8 -*-
The 17th International Conference on Head-Driven Phrase Structure Grammar (2010) was held at Université Paris Diderot, Paris 7.

The conference featured 2 invited talks and 19 papers, and 5 posters selected by the program committee (Doug Arnold,
Emily M. Bender, Philippe Blache, Olivier Bonami (chair), Bob Borsley, Gosse Bouma, Rui Chaves, Ann
Copestake,
Berthold Crysmann, Kordula De Kuthy, Dan Flickinger, Daniele Godard, Anke Holler, Jean-Pierre
Koenig, Valia Kordoni, Anna Kupsc, Bob Levine, Rob Malouf, Nurit Melnik, Philip Miller, Stefan
Müller, Gerald Penn, Frank Richter, Ivan Sag, Manfred Sailer, Jesse Tseng, Frank Van Eynde, Gert
Webelhuth, Shuichi Yatabe, Eun-Jung Yoo).

A workshop about \emph{Morphology and Formal Grammar}
was attached to the conference. It featured one invited talk and 10 papers and three posters, selected by the program
committee of this workshop (Farrell Ackerman,
Emily Bender,
James Blevins,
Olivier Bonami (chair),
Dunstan Brown,
Gilles Boyé,
Berthold Crysmann,
Bernard Fradin,
Rob Malouf,
Stefan Müller,
Louisa Sadler,
Pollet Samvelian,
Andrew Spencer,
Jesse Tseng,
Gert Webelhuth).


In total there were 29  submissions to the conference and 24 submissions to the workshop.
We want to thank the respective program committees for putting this nice program together.



Thanks go to Anne Abeillé (chair), Gabriela Bilbîie, Olivier Bonami, Marianne Desmets, Danièle Godard,
Fabiola Henri,
Frédéric Laurens,
Philip Miller,
François Mouret,
Clément Plancq,
Jana Strnadová,
Delphine Tribout,
Géraldine Walter,
Grégoire Winterstein, who were in charge of local arrangements.


As in the past years the contributions to the conference proceedings are based on the five page abstract
that was reviewed by the respective program committees, but there is no additional reviewing of the
longer contribution to the proceedings.
To ensure easy access and fast publication we have chosen an electronic format.


The proceedings include all the papers except those by Farrell Ackerman and Rob Malouf, Dan Flickinger, Jean-Pierre Koenig and Karin
Michelson, Robert Levine, Jakob Maché, Nurit Melnik, and Gregory Stump.
