%% -*- coding:utf-8 -*-
The Joint 2016 Conference on Head-driven Phrase Structure Grammar and
Lexical Functional Grammar (HeadLex) was held at the Polish Academy of
Sciences, Warsaw.  This conference conjoined the annual Conference on
Head-driven Phrase Structure (HPSG) and the annual Conference on
Lexical Functional Grammar (LFG) in that a joint program committee and
a joint organizing committee were instituted to plan and carry out a
joint conference, also resulting in a joint proceedings.

The conference featured 2 invited talks, 32 papers, and 10
posters and were selected by:  

Doug Arnold (program committee co-chair)
    Berthold Crysmann (program committee co-chair),
    John Lowe (program committee co-chair),
    Ida Toivonen (program committee co-chair),
Anne Abeillé,
    Alex Alsina,
    Avery Andrews,
    I Wayan Arka,
    Ash Asudeh,
    Dorothee Beermann,
    Oleg Belyaev,
    Emily M. Bender,
    Adams Bodomo,
    Tina Bögel,
    Olivier Bonami,
    Francis Bond,
    Kersti Borjars,
    Bob Borsley,
    George Aaron Broadwell,
    Miriam Butt,
    Rui Chaves,
    Philippa Cook,
    Ann Copestake,
    Mary Dalrymple,
    Elisabet Engdahl,
    Daniel Flickinger,
    Martin Forst,
    Hyun-Jong Hahm,
    Dag Haug,
    Fabiola Henri,
    Anke Holler,
    Ron Kaplan,
    Anna Kibort,
    Jong-Bok Kim,
    Tracy Holloway King,
    Jean-Pierre Koenig,
    Helge Lodrup,
    Nurit Melnik,
    Laura Michaelis,
    Stefan Müller,
    Louise Mycock,
    Tsuneko Nakazawa,
    Tatiana Nikitina,
    Rachel Nordlinger,
    Stephan Oepen,
    Ryo Otoguro,
    Agnieszka Patejuk,
    Gerald Penn,
    Adam Przepiórkowski,
    György Rákosi,
    Frank Richter,
    Victoria Rosén,
    Jeffrey Runner,
    Louisa Sadler,
    Manfred Sailer,
    Peter Sells,
    Liselotte Snijders,
    Andrew Spencer,
    Frank Van Eynde,
    Nigel Vincent,
    Stephen Wechsler,
    Shuichi Yatabe,
    Eun-Jung Yoo,
    Annie Zaenen.


% wie viele?
%In total there were x  submissions to the conference and x submissions to the workshop.
We want to thank the program committee for putting together a very
successful program.

Thanks go to Agnieszka Patejuk (co-organizer) and Adam Przepiórkowski
(co-organizer), who were in charge of local arrangements, and their
assistants Katarzyna Krasnowska-Kieraś, Jakub Kozakoszczak and Jan
Ziółkowski.

 Original conference submissions were five page abstracts. The revised versions of the papers
underwent an additional round of reviweing. We thank the anonymous reviewers for their help in reviewing.

To ensure easy access and fast publication we have chosen an
electronic format. This joint proceedings also functions as the annual
proceedings of each of the usually separate conferences.


The proceedings include all the papers except the ones by 
Liesbeth Augustinus,
Tina Bögel, 
Kersti Börjars and John Payne, 
Lionel Clément and Sekou Diao, 
Hyun Jong Hahm, 
Anna Kibort, 
Tibor Laczkó, 
Frank Van Eynde, 
Nigel Vincent and Kersti Börjars.


