%% -*- coding:utf-8 -*-
The 19th International Conference on Head-Driven Phrase Structure Grammar (2012) was held at the Chungnam National University Daejeon.

The conference featured 1 invited talk and 20 papers selected by the program committee (Anne
Abeillé,
Doug Arnold
Emily M. Bender,
Francis Bond,
Oliver Bonami,
Bob Borsley,
Rui Chaves,	Ann Copestake,
Berthold Crysmann, 	Elisabet Engdahl,
Dan Flickinger, 	Jong-Bok Kim (Chair),
Tibor Kiss, 	Jean-Pierre Koenig,
Valia Kordoni, 	Bob Levine,
Laura Michaelis, 	Stefan Müller,
Tsuneko Nakazawa, 	Jeff Runner,
Ivan Sag, 	Manfred Sailer,
Frank Van Eynde, 	Gert Webelhuth,
Eun-Jung Yoo, 	Stephen Wechsler, and
Shuichi Yatabe).

A workshop about \emph{Ellipsis and Formal Grammar}
was attached to the conference. It featured one invited talk and 7 papers and a poster, selected by the program
committee of this workshop (Doug Arnold, 	John Beavers,
Rui Chaves, 	Jong-Bok Kim (Chair),
Jason Merchant, 	Stefan Müller,
Myung-Kwan Park, 	Eric Potsdam,
Ivan Sag, 	Mark Steedman,
Tom Wasow, 	Shuichi Yatabe, and
Eun-Jung Yoo).

% wie viele?
%In total there were 29  submissions to the conference and 24 submissions to the workshop.
We want to thank the respective program committees for putting this nice program together.



Thanks go to Byong-Rae Ryu (chair) and Hee-Rahk Chae (Hankuk University of Foreign Studies),
Myong-Hi Chai (Chosun College of Science and Technology), Sae-Youn Cho (Kangwon National
University), Incheol Choi (Kyungpook National University), Chan Chung (Dongseo University), Munpyo
Hong (Sung\-kyun\-kwan University), Hanmin Jung (Korea Institute of Science and Technology Information),
Kil-Soo Ko (Seoul National University), Hae-Yun Lee (Hankuk University of Foreign Studies), Nam-Keun
Lee (Chosun University), Seungwoo Lee (Korea Institute of Science and Technology Information),
Yonghun Lee (Chungnam National University), Kyung-Sub Lim (Dongshin University), Yongkyoon No
(Chungnam National University), Hyopil Shin (Seoul National University), Won-Kyung Sung (Korea
Institute of Science and Technology Information), and Cheongmin Yook (Keimyung University), who were
in charge of local arrangements.
 
We also thank the conference chair, Byung-Soo Park (Kyunghee University, emeritus), and the members
of the Advisory Committee for the Local Organizing Committee, especially Suk-Jin Chang (Seoul
National University, emeritus), Kiyong Lee (Korea University, emeritus), Ik-Hwan Lee (Sangmyung
University), Chungmin Lee (Seoul National University, emeritus), Key-Sun Choi (KAIST), Jae-Woong
Choe (Korea University), Sang-Kyu Park (Electronics and Telecommunications Research Institute), and
Seungho Nam (Seoul National University).


As in the past years the contributions to the conference proceedings are based on the five page abstract
that was reviewed by the respective program committees, but there is no additional reviewing of the
longer contribution to the proceedings.
To ensure easy access and fast publication we have chosen an electronic format.

The proceedings include all the papers except those by Joshua Crowgey (the poster), David Erschler, Mark Steedman, and Takafumi Maekawa.
