%% -*- coding:utf-8 -*-
The 18th International Conference on Head-Driven Phrase Structure Grammar (2011) was held at the University of Washington.

The conference featured 2 invited talks and 16 papers, and 1 poster selected by the program committee (Anne Abeillé,
Doug Arnold,
Emily M. Bender,
Philippe Blache,
Olivier Bonami,
Robert Borsley,
Gosse Bouma,
Rui Chaves,
Berthold Crysmann (chair),
Dan Flickinger,
Danièle Godard,
Lars Hellan,
Anke Holler,
Jong-Bok Kim,
Jean-Pierre Koenig,
Valia Kordoni,
Anna Kupsc,
Robert Levine,
Rob Malouf,
Nurit Melnik,
Philip Miller,
Stefan Müller,
Gerald Penn,
Adam Przepiorkowski,
Frank Richter,
Ivan Sag,
Manfred Sailer,
Jesse Tseng,
Frank Van Eynde,
Gert Webelhuth,
Stephen Wechsler,
Shuichi Yatabe).

A workshop about \emph{Information Structure and Formal Grammar}
was attached to the conference. It featured one invited talk and 8 papers and a poster, selected by the program
committee of this workshop (Felix Bildhauer
Daniel Büring
Berthold Crysmann (chair)
Kordula De Kuthy
Elisabet Engdahl
Claire Gardent
Jonathan Ginzburg
Tracy Holloway King
Manfred Krifka
Jean-Marie Marandin
Laura Michaelis
Stefan Müller
Irina Nikolaeva
Patrizia Paggio
Arndt Riester
Mats Rooth
Mark Steedman
Malte Zimmermann).

%
% 31 submissions (mail 04.04.2011) 15 submissions to the workshop
We want to thank the respective program committees for putting this nice program together.



Thanks go to Emily M. Bender (chair), Joshua Crowgey,
Michael Goodman,
Varya Gracheva,
Prescott Klassen,
Naoko Komoto,
Clarissa Surek-Clark,
Emily Silgard,
Sanghoun Song,
Lisa Tittle, and David Wax, who were in charge of local arrangements.


As in the past years the contributions to the conference proceedings are based on the five page abstract
that was reviewed by the respective program committees, but there is no additional reviewing of the
longer contribution to the proceedings.
To ensure easy access and fast publication we have chosen an electronic format.


The proceedings include all the papers except those by Olivier Bonami, Rui Chaves, Anna Gazdik,
Tibor Kiss, Mats Rooth, and Thomas Wasow and David Clausen.
