%% -*- coding:utf-8 -*-
The 28th International Conference on Head-Driven Phrase Structure Grammar (2021) took place as an
online conference and was organized by Frank Richter (Goethe-Universität, Frankfurt am Main).


The conference featured 2 invited talks and 12 papers selected by the program committee 
(Anne Abeillé, 
    Emily M. Bender, 
    Felix Bildhauer, 
    Hans Boas, 
    Olivier Bonami, 
    Francis Bond, 
    Gosse Bouma, 
    Rui Chaves, 
    Berthold Crysmann, 
    Daniel Flickinger, 
    Fabiola Henri, 
    Thomas Hoffmann, 
    Anke Holler, 
    Gianina Iordachioaia, 
    Paul Kay, 
    Jong-Bok Kim, 
    Jean-Pierre Koenig, 
    David Lahm, 
    Nurit Melnik (chair),
    Laura Michaelis, 
    Philip Miller, 
    Stefan Müller, 
    Tsuneko Nakazawa, 
    Petya Osenova, 
    Rainer Osswald, 
    Gerald Penn, 
    Frank Richter, 
    Manfred Sailer, 
    Stephen Wechsler, 
    Shuichi Yatabe,
    Eun-Jung Yoo, 
    Olga Zamaraeva). There was a workshop on negation with one invited speaker and six regular papers.


% wie viele?
%In total there were x  submissions to the conference and x submissions to the workshop.
We want to thank the program committees for putting these nice programs together.

% Thanks go to Gabriela Bîlbîie and Emil Ionescu, who were in charge of local
% arrangements.
 

As in the past years the contributions to the conference proceedings are based on the five page abstract
that was reviewed by the respective 
program committee, but there is no additional reviewing of the
longer contribution to the proceedings. To ensure easy access and fast publication we have chosen an
electronic format. As of this year, the proceedings will be published by the University Library of
Goethe-Universität, Frankfurt am Main. Previous volumes that were published by CSLI Publications
have been migrated to the new location. We want to thank CSLI Publications for the good
collaboration in all these years.

The proceedings include all the papers of the conference and workshop except the ones by Gabriel
Aguila-Multner \& Berthold Crysmann, Zahra Mirrazi \& Hedde Zeijlstra.


