%% -*- coding:utf-8 -*-
The 15th International Conference on Head-Driven Phrase Structure Grammar (2008) was held in Keihanna, Japan
and organized by the National Institute of Information and Communications Technology and the Shoin Institute for Linguistic Sciences,
Shoin Women's University.

The conference featured 2 invited talks and 17 papers
selected by the program committee 
(Anne Abeillé,
Olivier Bonami,
Francis Bond,
Bob Borsley,
Gosse Bouma,
Rui Chaves,
Berthold Crysmann,
Markus Egg,
Elisabet Engdahl,
Dan Flickinger,
Jonathan Ginzburg,
Danièle Godard,
Takao Gunji,
Chikara Hashimoto,
Erhard Hinrichs,
Anke Holler,
Chiharu Uda Kikuta,
Jong-Bok Kim,
Tibor Kiss,
Anna Kupsc,
Shalom Lappin,
Robert Levine,
Rob Malouf,
Nurit Melnik,
Detmar Meurers,
Stefan Müller,
Tsuneko Nakazawa,
Gerald Penn,
Adam Przepiórkowski,
Frank Richter,
Louisa Sadler,
Ivan Sag,
Manfred Sailer,
Jesse Tseng,
Stephen Wechsler,
Sh{\^{u}}ichi Yatabe (chair),
Kei Yoshimoto).

A workshop about \emph{Grammar at the Interfaces}
was attached to the conference. It featured one invited talk and 7 papers, selected by the program committee.

In total there were 34 submissions to the main conference and to the
workshop. 
We want to thank the respective program committee for putting this nice program together.



Thanks go to Francis Bond, Takao Gunji, Sanae Fujita, Kyoko Kanzaki, Taka\-yuki Kuribayashi, and Eric
Nichols, who were in charge of local arrangements.


As in the past years the contributions to the conference proceedings are based on the five page abstract
that was reviewed by the respective program committees, but there is no additional reviewing of the
longer contribution to the proceedings.
To ensure easy access and fast publication we have chosen an electronic format.


The proceedings include all the papers except those by Stefan Müller and Tsuneko Nakazawa.

