%% -*- coding:utf-8 -*-
The 20th International Conference on Head-Driven Phrase Structure Grammar (2013) was held at the
Freie Universität Berlin.

The conference featured 2 invited talks and 15 papers selected by the program committee 
(Emily M. Bender,
    Olivier Bonami,
    Bob Borsley,
    Rui Chaves,
    Berthold Crysmann,
    Kordula De Kuthy,
    Elisabet Engdahl,
    Daniel Flickinger,
    Jong-Bok Kim,
    Jean-Pierre Koenig,
    Valia Kordoni,
    Anna Kupsc,
    Robert Levine,
    Nurit Melnik,
    Stefan Müller,
    Tsuneko Nakazawa,
    Gerald Penn,
    Adam Przepiórkowski,
    Frank Richter,
    Louisa Sadler,
    Ivan Sag,
    Manfred Sailer,
    Jesse Tseng (chair),
    Frank Van Eynde,
    Gert Webelhuth,
    Stephen Wechsler,
    Shuichi Yatabe,
    Eun-Jung Yoo).

A tutorial \emph{Linguistic Research with Large Annotated Web Corpora} by Felix Bild\-hauer
and Roland Schäfer and a workshop about \emph{Progress in Linguistics}
were attached to the conference. The workshop program was put together by Stefan Müller and consisted of
invited talks only.

% wie viele?
%In total there were 27  submissions to the conference and 0 submissions to the workshop.
We want to thank the respective program committees for putting this nice program together.

Thanks go to Stefan Müller (chair), Viola Auermann, Lea Helmers, and Jakob Maché, who were
in charge of local arrangements.
 
The conference was supported by a grant from the Deutsche Forschungsgemeinschaft to Stefan Müller (MU 2822/7-1).

As in the past years the contributions to the conference proceedings are based on the five page abstract
that was reviewed by the respective program committees, but there is no additional reviewing of the
longer contribution to the proceedings.
To ensure easy access and fast publication we have chosen an electronic format.

The proceedings include all the papers except those by  Farell Ackerman, Rob Malouf, and John Moore,
Tibor Kiss, Takafumi Maekawa, Gereon Müller, Stefan Müller (workshop), Andreas Pankau, Frank
Richter, Anatol Stefanowitsch, and Nigel Vincent.

