%% -*- coding:utf-8 -*-
The 10th International Conference on Head-driven Phrase Structure Grammar was held at
Michigan State University, Michigan in the USA. 

The conference featured three invited talks, 26 papers, and two alternate papers,
selected by the program committee 
(Bob Borsley, chair,
Doug Arnold,
Elisabet Engdahl,
Erhard Hinrichs,
Tom Hukari,
Andreas Kathol,
Jean-Pierre Koenig,
Shalom Lappin,
Detmar Meurers,
Tsuneko Nakazawa,
Adam Przepiórkowski,
Ivan Sag,
Gert Webelhuth,
Shûichi Yatabe) with the help of additional reviewers (Ronnie Cann,
Danièle Godard, 
Georgia Green,
Jeanette Gundel,
Caroline Heycock,
Ewan Klein,
Stefan Müller,
Frank Richter,
Manfred Sailer, 
Andrew Spencer).
In total there were 42 submissions.
We want to thank the program committee and the external reviewers
for putting this nice conference program together.

Thanks go to Ivan Sag and Gert Webelhuth, who were in charge of local arrangements.
I also want to thank Ivan Sag for help regarding computational infrastructure in Stanford.

As was decided on the Business Meeting I will take care of editing the HPSG proceedings from now on,
in order to guarantee fast publication of the conference results
to make the work presented at the conference available to a wider audience. 
As in the past years the contributions to the conference proceedings are based on the five page abstract
that was reviewed by the program committee, but there is no additional reviewing of the
longer contribution to the proceedings.
To ensure easy access and fast publication we have chosen an electronic format.

The proceedings include all the papers except those by 
Olivier Bonami and Danièle Godard,
Gosse Bouma,
Incheol Choi and Stephen Wechsler,
Robert Malouf,
Vanessa Metcalf,
Peter Sells,
and Kei Yoshimoto and Masahiro Kobayashi.
