\documentclass[twoside]{book}


\usepackage[FINAL]{cslipubs} % CSLI Pubs format; use [FINAL] for final copy.
\usepackage[comma]{natbib} % CSLI Pubs favored bibliography package.
\usepackage{chapterbib} % Allows chapters to have separate bibliographies.

\title{The Proceedings of the 9th International Conference on HPSG} % \title{__} is necessary.

\author{Jong-Bok Kim and Stephen Wechsler}% \author{__} is necessary.

\usepackage{lingmacs}

\usepackage[comma]{natbib} % CSLI Pubs favored bibliography package.
\usepackage{avm,ch-lingmacros,tree-dvips}
\usepackage{import}         % Allows chapters to have separate subdirectories.
\usepackage{chapterbib}     % Allows chapters to have separate bibliographies.


% Hankuk Univ. of
%Foreign Studies hrchae@hufs.ac.kr}
%\date{ }


\def\boxit#1{\lower1.3pt\hbox{\hspace*{1pt}\vbox{\hrule height.3pt
        \hbox{\vrule width.3pt\kern1.0pt\vbox{\kern1.0pt
        \hbox{\rm #1}\kern1.5pt}\kern1.0pt\vrule width.3pt}
                \hrule height.3pt}\hspace*{1pt}}}

\def\tr{\underbar{\ \ \ }\ }



\setcounter{chapter}{2}
 \setcounter{page}{43}

\begin{document}
\achapter{Differences between Externally and Internally Headed
Relative Clause Constructions}{Chan Chung and Jong-Bok Kim}
\markright{Externally and Internally Headed Relative Clauses}

\section{Introduction}
In terms of truth conditional meanings, there is no clear
difference between (Korean) IHRCs (internally head relative) like
\ex{1}a and EHRCs (externally headed relative) like \ex{1}b.

\eenumsentence{
\item\shortexnt{6}
{Tom-un  &[sakwa-ka& cayngpan-wi-ey& iss-nun &kes]-ul} {Tom-{\sc
top} & apple-{\sc nom}&tray-{\sc top}-{\sc loc} &exist-{\sc pne} &
{\sc kes}-{\sc acc}}

\shortex{2}  {mekessta.} {ate} {`Tom ate an apple, which was on
the tray.'}

\item\shortex{7}
{Tom-un  &[\tr & cayngpan-wi-ey & iss-nun & sakwa]-ul& mekessta.}
{Tom-{\sc top} &     &tray-{\sc top}-{\sc loc} &exist-{\sc pne}
&apple-{\sc acc} &ate} {`Tom ate an apple that was on the tray.'}}


\noindent They all describe events of an apple's being on the tray
and Tom's eating it. But, there exist several intriguing
differences between the two constructions.
% which no previous
%approaches have seriously attempted to account for.
One crucial difference between the IHRC and EHRC comes from the
fact that  the semantic object of {\it mekessta} `ate' in IHRC
examples like \ex{0}a is the NP {\it sakwa} `apple' buried inside
the embedded clause followed by {\it kes}. It is thus the subject
of the embedded clause {\it apples} that serves as the semantic
argument of the main predicate.

In the analysis of such IHRCs, of central interest are thus (a)
how we can analyze the constructions in syntax and (b) how we can
associate the internal head of the IHRC clause with the matrix
predicate so that the head can function as its semantic argument,
and (c) what makes the differences between the two constructions.
This paper is an attempt to provide answers to such recurring
questions within the framework of HPSG.



\section{Syntax of the IHRC}

\subsection{Internal Syntax}

%The first thing we need to figure out in the IHRC is the internal
%properties of the clause and {\it kes}, in addition to their
%relationship.

One noticeable morphological property of the IHRC construction is
that as shown in \ex{0}a, the embedded clausal predicate should be
in the adnominal form of {\it (n)un}, followed by {\it
kes}.\footnote{The restriction on the types of adnominal form
seems to be related to semantic and pragmatic constraints. See
section 4.} This is another main difference from EHRC, in that the
predicate in EHRCs can have three different types of suffixes
sensitive to tense information. This contrast is given in
\ex{1}:\footnote{These three basic kinds of tense-sensitive
prenominal markers in the EHRC can be extended to denote aspects
when combined with tense suffixes. Thus the possible prenominal
verb forms are {\it ilk-ten} `read-progressive', {\it ilk-essten}
`read-past progressive', {\it ilk-essul} `read-past conjecture',
{\it ilk-essessul} `read-past perfective conjecture', {\it ilk-ko
issten} `past perfective progressive'}

\eenumsentence{
\item\shortex{4}
{Tom-i    &\tr$_{i}$ &ilk-nun/un/ul &chayk$_{i}$} {Tom-{\sc nom} &
&read-{\sc pres.pne/pst.pne/fut.pne}    & book} {`the book that
Tom reads'}
\item\shortexnt{6}
{Tom-un  &[sakwa-ka& cayngpan-wi-ey& {\bf iss-nun/*ul} &kes]-ul}
{Tom-{\sc top} & apple-{\sc nom}&tray-{\sc top}-{\sc loc}
&exist-{\sc pne}  &{\sc kes}-{\sc acc}}

\shortex{2} {mekessta.}
 {ate} {`Tom ate an
apple, which was on the tray.'}}

\noindent
%As given in the contrast, unlike the EHRC, the IHRC
%chan does not allow a future %chan prenominal marker {\it %chan (n)un %chan}.
%chan Change footnote 1to footnote 2 and footnote 2 to footnote 1 %chan

In the traditional Korean grammar, {\it kes} in the IHRC is called
a `dependent noun' in that it always requires a verb of an
adnominal form and cannot exist alone as a word:

\enumsentence{ \shortexnt{7} {(Na-nun  &    totwuk-i &
unhayng-eyse)& *(nao-nun) & kes-ul}  {I-{\sc top} & thief-{\sc
nom} & bank-from &come-out-{\sc pne} &{\sc kes}-{\sc acc}}

\shortexnt{2}{capassta.} {caught}}

\noindent Example \ex{0} shows that the adnominal verb {\it
nao-nun} is neither optional nor can be realized as an empty
category.

A tight syntactic relation between the clausal predicate and the
noun {\it kes} can also be found from the optionality of the IHRC
in \ex{1}:

\eenumsentence{ \item\shortexnt{8} {Na-nun  &  *(kangto-ka &
unhayng-eyse & nao-nun)&          kes-ul} { I-{\sc top} &
robber-{\sc nom} & bank-from     & come-out-{\sc pne} & {\sc
kes-acc}}

\shortex{2} {capassta.} {caught} {`I arrested the robber who was
coming out of the bank.'}

\item\shortex{8} {Na-nun  &  (unhayng-eyse & nao-nun)& kangto-ul &
capassta.} { I-{\sc top}  &  bank-from & come-out-{\sc pne}
&kangto-{\sc acc} & caught} {`I arrested the robber who was coming
out of the bank.'}}

\noindent The IHRC example in \ex{0}a indicates that the adnominal
IHRC clause as well as its predicate is an obligatory element.
However, the entire EHRC clause in \ex{0}b is optional.

The point to note is that in canonical control constructions the
matrix verb can exist as an independent word, without the governed
verb as in \ex{1}a. This is different from a canonical complex
predicate construction as in \ex{1}b:

\eenumsentence{\item\shortex{8} {(Na-nun  &  John-hanthey &
sakwa-lul & mek-ulako) & seltukhayssta.}{I-{\sc top}   & John-{\sc
dat}& apple-{\sc acc} &eat-{\sc pne} &persuaded} {`I persuaded
John to eat an apple.'}
\item\shortex{8}
{na-nun & sakwa-lul & mek-e & poassta} {I-{\sc top} & apple-{\sc
acc} & eat-{\sc comp} & tried}{`I tried to eat an apple.'}}

\noindent Such observations support the assumption that the
pre-adnominal verb and {\it kes} forms a syntactic unit, possibly
functioning as a complex predicate, as argued by Chung (1999).

There seems to exist additional phenomena showing the parallelism
between the IHRC and verbal complex constructions. One such
phenomenon is the so-called afterthought expression construction:

\eenumsentence{{\small
\item\shortex{8}{*Na-nun &    kes-ul &   capassta, &
totwuk-i & unhayng-eyse  &nao-nun.} {I-{\sc top}   & {\sc
kes}-{\sc acc} & arrested & thief-{\sc nom}  & bank-from  &
come-out-{\sc pne}} {`I arrested the thief who was coming out of
the bank.'}

\item\shortexnt{8}
{na-nun   & totwuk-ul & capassta, & unhayng-eyse  &nao-nun.} {
I-{\sc top} & thief-{\sc acc}  & arrested  & bank-from &
come-out-{\sc pne}}}}

\noindent Sentence \ex{0}a is an instance of the IHRC
construction, where the IHRC is used as an afterthought
expression. It shows that the IHRC cannot be used as an
afterthought expression differently from the EHRC in \ex{0}b.
 It suggests that {\it kes} and the adnominal verb constitute a
syntactic unit and that they cannot be separated. The same pattern
is also observed in the verbal complex construction:

\eenumsentence{
\item\shortex{8}
{ *Na-nun    & poassta,  & sakwa-lul  &  mek-e.}
       {I-{\sc top}  &      tried  &   apple-{\sc acc} & eat}
       {`I tried an apple.'}
\item\shortex{8}
 {Na-nun  &  seltukhayssta, & John-hanthey & sakwa-lul &
 mekulako.}
      {I-{\sc top} &      pesuaded   &   John-{\sc dat} &          apple-{\sc acc}
        &
      eat}
      {`I persuaded John to eat an apple.'}}

\noindent \ex{0}a is an instance of the verbal complex where the
auxiliary verb and its governed verb cannot be separated, while
\ex{0}b is an instance of the control verb construction where the
matrix verb and its complement are separable.

Besides the arguments based on the lexical integrity, another
parallelism between the IHRC and verbal complex constructions
arises from the fact that the heads of the constructions, namely,
{\it kes} in the IHRC and the auxiliary verb in the verbal
complex, are a kind of clitics, diachronically derived from
independent words whose phonetic forms are the same. For example,
the auxiliary verb {\it pota} `try as a test' in \ex{-2}b and
\ex{0}a, has a non-auxiliary-verb counterpart {\it pota} `see',
which can be used as an independent word. The same observation can
be made in the IHRC. The head {\it kes} in the IHRC can never be
used as a referring expression and never takes a specifier such as
{\it ku} `the' and {\it ce} 'that':

\eenumsentence{
\item\shortexnt{8}
{*Na-nun  &   totwuk-i  &  unhayng-eyse & nao-nun & ku}
 {I-{\sc top}     &  thief-{\sc nom} &  bank-from &
come-out-{\sc pne} & the}

\shortex{2}
 { kes-ul & capassta.}
 { {\sc kes}-{\sc acc} & caught} {`I
arrested the thief who was coming out of the bank.'}}

However, there exists a referential noun counterpart kes `thing',
which can be used as a referring expression and can take a
specifier:


\enumsentence{ \shortex{8} {Na-nun & ku   &   kes-ul & sassta.} {
I-{\sc top} & the & thing-{\sc acc} & bought}{`I bought the thing
(it).'}}


To sum up, there are some parallelisms between the verbal complex
and the combination of ``adnominal verb + kes" in the IHRC phrase.
It suggests that the combination in the IHRC needs to be treated
as a syntactic unit, namely, as a complex noun.
%It also suggests
%that the adnominal verb needs to be considered to have a dual
%function of a verbal complement and an adjunct.


% Example \ex{1} shows that there could be no
%intervening elements between the clausal predicate and {\it kes}
%(cf. Chung 1999):
%
%\enumsentence{ \shortexnt{4} {*kyongchal-i & [{\bf kangto-ka} &
%{\bf unhayng-eyse}  & {\bf nao-nun}]} { police-{\sc nom}&
%[theif-{\sc nom}& bank-from     & come.out-{\sc pne}]}%
%
%\shortex{4} {ecey & {\bf kes}-ul & chephohayssta} {yesterday  &
%{\sc kes}-{\sc acc} & arrested} {`(int.) The police arrested a
%thief who was coming out of the bank yesterday.'}}
%
%In a similar respect,

The contrast in \ex{1} shows that whereas more than one EHRC
clause can be stacked together, only one IHRC clause is possible:

\eenumsentence{\item\shortexnt{6} {*kyongchal-i & [{\bf kangto-ka}
& {\bf unhayng-eyse}  & {\bf nao-nun}]} {police-{\sc nom}&
[robber-{\sc nom}& bank-from     & come.out-{\sc pne}]}

\shortex{4} {[ton-ul & hwumchi-in] & {\bf kes}-ul & chephohayssta}
{money-{\sc acc} & steal-{\sc pne} & {\sc kes}-{\sc acc} &
arrested} {`(int.) The police arrested a thief coming out of the
bank, stealing money.'}

\item\shortexnt{5} { kyongchal-i & [\tr & {\bf unhayng-eyse} & {\bf
nao-nun}]} { police-{\sc nom}& [& bank-from & come.out-{\sc pne}]}

\shortex{4} {[ton-ul & hwumchi-in] & {\bf kangto}-lul &
chephohayssta} {money-{\sc acc} & steal-{\sc pne} & robber-{\sc
acc}-{\sc acc} & arrested} {`(int.) The police arrested a thief
coming out of the bank, stealing money.'}}

\noindent This contrast implies that the adnominal clause in the
IHRC has the canonical properties of a complement clause.

Given these observations showing a strong syntactic bondage
between {\it kes} and the adnominal verb give us enough reason to
take the verb {\it -kes} as a complex element as represented in
the following lexical entry:

\enumsentence{Lexical Entry for {\it kes} (first approximation):\\
\begin{avm}\avml
\[\q<kes\q>\\
  HEAD \; {\it noun} \\
  ARG-ST \<V\[FORM {\it (n)un}\\
              ARG-ST \@a\]\> $\oplus$ \@a\]
 \avmr
\end{avm}
}

\noindent The lexical entry in \ex{0} specifies that the {\it kes}
noun selects as its argument a verbal element as well as the
arguments that this verb selects.
%\footnote{Since the SUBJ value is
%empty, the S[VFORM {\it nun}] in the ARG-ST will be mapped onto
%the COMPS. See Sag and Wasow (1999) for how arguments are realized
%into grammatical functions.}
The argument selection requirements
of the adnominal verb are thus passed to the head {\it kes} with
which it combines. This lexical information in turn means that the
IHRC will have the following internal structure:\footnote{Adopting
Bratt (1995), we assume that Korean allows two lexical elements to
combine to form a subphrasal element:

\enumsentence[(i)]{Lexical Head-Complement Schema:\\
X$'$ $\rightarrow$ Comp[$+$LEX], H[$+$LEX]}

\noindent This schema captures the constituenthood of the
preceding main verb and the following auxiliary verb. See Sells
1995 and Chung 1998 for a similar analysis.}


 \enumsentence{{\small

\begin{tabular}[t]{cccccccccc}
                &\node{a}{NP}\\[2ex]
  \node{b}{...}  &                      &\node{c}{N} \\[2ex]

                 &\node{d}{V[VFORM (n)un]}&&\node{e}{N}\\[2ex]

        &\node{g}{...}&  &\node{h}{kes}
\end{tabular}
\nodeconnect{a}{b}\nodeconnect{a}{c}
\nodeconnect{c}{d}\nodeconnect{c}{e}
\nodeconnect{d}{g}\nodeconnect{e}{h} } }


\noindent Such a structure, combined with the other universal
constraints of the HPSG, will generate the following structure:

 \enumsentence{{\small {\smalltree{
\node{a}{NP}\\
\node{b}{\boxit{1}NP}&\node{c}{\begin{avm}\avml \hfil N$'$ \\[-1ex]
                                   \[SUBJ \q<\@1\q> \]\avmr\end{avm}}\\
                                   \\
\node{1}{sakwa-ka}&\node{d}{\boxit{2}NP}&\node{e}{\begin{avm}\avml
\hfil N$'$\\[-1ex]
          \[HEAD  verb\\
            SUBJ \q<\@1\q>\\
            COMPS \q<\@2\q> \] \avmr\end{avm}}\\
            \\

&\node{2}{cayngpan-wi-ey}

&\node{f}{{\small \begin{avm}\avml \hfil
\@3V\\[-1ex]
               \[%SUBJ  \q<\@1\q>\\
                 %COMPS \q<\@2\q>\\
                 ARG-ST \q<\@1,\@2\q> \]
\avmr\end{avm}}}&\node{g}{{\small \begin{avm}\avml \hfil N\\[-1ex]
               \[ARG-ST  \q<\@3\q> $\oplus$
                 \q<\@1,\@2\q>\]
\avmr\end{avm}}}\\
\\
&&\node{3}{iss-nun}&\hspace{0.5in}\node{4}{kes} \nodeconnect{a}{b}
\nodeconnect{a}{c} \nodeconnect{c}{d} \nodeconnect{c}{e}
\nodeconnect{e}{f} \nodeconnect{e}{g} \nodeconnect{b}{1}
\nodeconnect{d}{2}\nodeconnect{f}{3} \nodeconnect{g}{4}}}}}


\noindent The verb \ital{iss-nun} takes a subject and an oblique
complement. According to the lexical entry given in \ex{-2}, the
{\it kes} selects this verb as well as arguments via the argument
composition mechanism (indicated by $\oplus$). When the {\it kes}
 combines with the verb {\it iss-nun}, the result still
requires its oblique complement.  The resulting complex combines
with the oblique complement, forming a nominal phrase which in
turn combines with the subject NP. We thus eventually can see here
that the precise lexical information of the {\it kes} in the IHRC
projects a fully saturated nominal phrase. In what follows, we
will further see the nominal properties of the IHRC in its
external syntax.





\subsection{External Syntax}


Given the internal syntax of the IHRC, let us see the relationship
between the whole IHRC clause including {\it kes} and the matrix
verb.

To relate the matrix verb with this `internal semantic head', the
traditional transformational grammar has introduced empty
categories. For example, Ito (1985), Watanabe (1992), and Hoshi
(1996) introduce an empty head noun approach for Japanese IHRCs.
If interpreted for Korean,  the structure would be something like
the ones given in \ex{1}: \footnote{In Ito (1985) and Jhang
(1994), Japanese {\it no} and Korean {\it kes} is taken to be a
complementizer. }

\enumsentence{

\begin{tabular}[t]{cccccccccc}

a.  Ito (1985):            &&\node{1}{NP} \\
                \\

                &\node{2}{NP}  &             &\node{3}{NP}\\
\\
\node{4}{S}&     & \node{5}{N} &    \node{7}{e}
\\
\\
 \node{a}{...}&        & \node{6}{kes}

\end{tabular}
\nodeconnect{1}{2}\nodeconnect{1}{3}\nodeconnect{2}{4}
\nodeconnect{2}{5}\nodeconnect{5}{6}\nodeconnect{3}{7}
\nodetriangle{4}{a}
\bigskip

\begin{tabular}[t]{cccccccccc}

b. Hoshi (1996):            &     &\node{1}{NP} \\
                \\

                &\node{2}{NP-case}  &             &\node{3}{NP}\\
\\
\node{4}{IP}        &\node{5}{N} &            &\node{6}{pro}
\\
\\
\node{7}{...NP$_{i}$...}&\node{8}{kes}
\end{tabular}
\nodeconnect{1}{2}\nodeconnect{1}{3}\nodeconnect{2}{4}
\nodeconnect{2}{5}\nodeconnect{4}{7}\nodeconnect{5}{8}
\nodeconnect{3}{6}}



\noindent Within Ito's head movement analysis, the internal head
NP in the embedded sentence moves into the head position in LF.
Hoshi's (1996) analysis posits the empty element {\it pro} is
adjoined to an NP headed by {\it kes} which is modified by an IP.
The {\it pro} is then coindexed with an NP within the IP.

Another direction that movement approaches have taken is to posit
an empty head or a {\it pro} to the right of the subordinate
clause and take the IHRC as an adjunct clause (Murasugi 1994, D.H.
Chung 1996). In particular, Chung (1996) introduces a null
perception predicate PRED, as represented in \ex{1}:

\enumsentence{\shortexnt{8} {Tom-un  &[sakwa-ka& cayngpan-wi-ey&
iss-nun &kes]-ul& PRED} {Tom-{\sc top} & apple-{\sc nom}&tray-{\sc
top}-{\sc loc} &exist-{\sc pne}  & {\sc kes}-{\sc acc}}

\shortex{2} {{\it pro} &  mekessta.} {  &ate}{`Tom ate the apples,
which were on the tray.'}}

\noindent His analysis allows the empty PRED to be interpreted as
predicates like {\it know, see, realize}, etc. The empty {\it pro}
is presumably bound by its antecedent in the clause.\footnote{One
of the arguments for the adjunct clause lies on the fact that {\it
kes} can be replaced by a temporal element such as {\it swunkan}
`moment', {\it hyoncang} `place'.}

Leaving aside the evaluation of such traditional empty-head
analyses in detail, we claim that we could capture various
properties of the construction even without positing empty
elements, and further that there exist ample evidence supporting
that the construction is a direct syntactic nominal complement of
the matrix predicate.\footnote{See Y.B. Kim (2002) for detailed
criticisms against an adjunct analysis of Chung (1996)}


A strong argument against an adjunct treatment centers on the
passivization of the IHRC clause. As in \ex{1}, the object IHRC
clause can be promoted to the subject of the sentence.

\enumsentence{ \shortex{5} {[Tom-i & talli-nun &  kes]-i  &
Mary-eyeuyhayse & caphiessta} {Tom-{\sc nom}& run-{\sc pne}  &
{\sc kes}-{\sc nom} & Mary-by & caught} {`Tom, who was running,
was caught by Mary.'} }

\noindent If we assume the IHRC clause is an adjunct clause from
semantic or syntactic reasons, we would then need to introduce a
system that an adjunct clause can participate in the passivization
process, contrary to most current practice. In contrast, the
present analysis where the IHRC clause is a nominal element would
not block the clause from being promoted to the subject from the
object.

A related problem of such an empty PRED approach would be that the
empty PRED cannot assign nominative case to subject IHRC phrases
like \ex{0} and \ex{1} since perception verbs such as {\it
realize}, {\it see}, etc. assign accusative case to its
complement. The case value is purely due to the main predicate
{\it salaciessta}:

\enumsentence{\shortexnt{6} {[sakwa-ka & cayngpan &wi-ey& iss-ten&
kes]-i& PRED]} {apple-{\sc nom} &tray& on-{\sc loc} & exist-{\sc
pst}-{\sc pne} & {\sc kes}-{\sc nom} & }

\shortex{5}{[{\it pro}  &salaciessta]} { &disappeared}{`The apple,
which was on the tray, disappeared.'}}

In addition, if there is an empty {\it pro} in the sentence with
the IHRC, there appears to be no reason to block us from replacing
it from an overt pronoun. But such a replacement is not possible:


\enumsentence{\shortexnt{6} {*[sakwa-ka & cayngpan &wi-ey&
iss-ten& kes-i& PRED]} {apple-{\sc nom} &tray & on-{\sc loc} &
exist-{\sc pst}-{\sc pne} &{\sc kes-nom} & }

\shortex{5}{[{\it ku kes-i}  &salaciessta]} {that one-{\sc nom}
&disappeared}{`The apple, which was on the tray, disappeared.'}}


In addition, the case marking pattern also implies that the
construction is a complement. Unlike the so-called complementizer,
{\it -ko}, the word {\it kes} can attract the nominative and
accusative case
 markings  as we have seen above. Some more examples are shown in
\ex{1} where {\it kes} attracts genitive or even instrument case markings:

%chan Replace (15a) with the example in footnote 6 below. (15a) is not a
%typical example of the IHRC since it has an individual level predicate.
%chan \footnote{It appears that a genitive case
%is marginally %chan acceptable. %chan
% chan
%{Meri-ka & eli-n & kes-i & phiano-lul & cal & chinta} {Mary-NOM &
%young-PNE & KES-NOM & piano-ACC & well & play} {`Mary, who is
%young, plays piano very well.'}
%chan
%Change this paragraph and the following examples as mentioned above % chan

\eenumsentence{
\item\shortex{8}
{?[Kangto-ka & unhayng-eyse & nao-nun & kes-uy] & chepho}
{robber-{\sc nom}  &bank-from& come-out-{\sc pne} & {\sc kes-gen}
& arrest}{ `the arrest of the robber who was coming out of the
bank.'}

\item\shortex{6} {[Mary-ka & ton-ul & pill-in &
kes]-{\bf ulo}  & chayk-ul & sassta} {Mary-{\sc nom}& money-{\sc
acc} & lend-{\sc pne} & {\sc kes}-{\sc inst} & book-{\sc acc} &
bought} {`Mary lent some money and bought a book with it.'}}

\noindent This case marking pattern shows that the IHRC is a
nominal projection. Such a case assignment pattern  is a canonical
property of a complement, rather than an adjunct.

%chan I think this paragraph is not necessary.
%
%Though there exist no definite criteria to distinguish complement
%from adjunct, as we have noted earlier, the possibility of
%passivizing the IHRC clearly indicates that the construction is a
%direct complement of the matrix predicate.
%chan

Based on these observations, we assume the structure \ex{0} for
the external structure of the IHRC.


\enumsentence{
\begin{tabular}[t]{cccccccccc}
           &                    \multicolumn{2}{c}{\node{1}{ VP}} \\[2ex]
               & \node{2}{\begin{avm}\avml \hfil NP$_{i}$\avmr\end{avm}}
          &
                           && \node{3}{V}\\[1ex]
\\[1ex]


       & \node{5}{\begin{avm}\avml \hfil .... N$_{i}$\\
        %\[RESTR \q<\@1\q>\]
        \avmr\end{avm}}
       & &&\node{6}{.....}\\[1ex]
\\[1ex]
\node{7}{}&\node{8}{kes}


\end{tabular}
\nodeconnect{1}{2}\nodeconnect{1}{3} \nodeconnect{3}{6}
\nodetriangle{2}{5}\nodeconnect{5}{8}}
%\nodeconnect{5}{9}
%\nodeconnect{7}{10}\nodeconnect{8}{11}}}


\section{Semantic Aspects of the IHRC}

%\subsection{The Function of {\it kes}}

Given the internal and external syntax of the IHRC, the remaining
issue is how to associate one of the arguments in the IHRC clause
as the semantic argument of the matrix predicate. As hinted
earlier, the approach we take is to assume that {\it kes} is a
kind of pronoun looking for its antecedent within the adnominal's
arguments as represented in \ex{1}:

\enumsentence{\label{entry2}Lexical Entry for {\it kes} (second
approximation):\\
\begin{avm}\avml
\[\q<kes\q>\\
  HEAD \; {\it noun} \\
  ARG-ST \@b\q<...[\;]$_{i}$...\q>\\
  CONTENT\|INDEX i \]\\

{\footnotesize (where \@b results from the argument composition
\<V\[FORM (n)un\\
  ARG-ST \@a\]\> $\oplus$ \@a)}
\avmr
\end{avm}
}

\noindent What this lexical entry tells us is that the index value
of {\it kes} is identical with either the adnominal verb or one of
the arguments that the adnominal verb selects (this plays an
important role in capturing an event as well as an entity
reading). The treatment of {\it kes} as a kind of pronoun gets
strong support from its pronominal properties (cf.\ D.H. Chung
1996). For example, the target of the internal head in \ex{0} is
highly dependent upon context:

\enumsentence{ \shortexnt{8} {[koyangi-ka & cwui-lul &ccoc-ko&
iss-nun & kes-ul]} {cat-{\sc nom}&mouse-{\sc acc} &chase-{\sc
comp} &in.state-{\sc pne}& {\sc kes}-{\sc acc}}

\shortex{2}{capassta}{caught}
{`(He) caught the mouse that the cat was chasing.\\
(He) caught the cat that was chasing the mouse.'}}

\noindent Depending on the context, the internal head could be
either the cat or the mouse or even both. Also, in the IHRC, {\it
kes} can have split antecedents as illustrated in \ex{1}.

\enumsentence{ \shortex{5} {[koyangi-ka & cwui-lul &ccoc-nun&
kes-ul]& katwuessta.} {cat-{\sc nom}&mouse-{\sc acc} &chase-{\sc
pne} &{\sc kes}-{\sc acc}& penned} {`(I) penned a cat chasing a
mouse.'}}

\noindent The target of the verb {\it katwuessta} `pen' could be
both `cat' and `mouse'.\footnote{Like \ex{-1}, the target could be
either {\it cat} or {\it mouse} too.}

 In addition, the {\it kes} in the construction can even have an
implicit antecedent, which is one of the canonical properties of
pronouns:


\enumsentence{\shortex{5} { [[khep-uy & mwul-i & nemchi-n]  &
kes-ul]& ttakassta.}
 {cup-{\sc gen} &water-{\sc nom}&overflowed-{\sc pne}& {\sc kes}-{\sc acc}
 &wiped-out}
 {`(I) wiped out the water that overflowed from the one in the
 cup.'}}

\noindent The interpretation we have for the example \ex{0} is
such that what I wiped out isn't the water in the cup but the one
that overflowed. There is no overt antecedent for the pronoun {\it
kes}.


Another point to note here is that IHRCs are syntactically very
similar to clausal complements. IHRCs and clausal complements both
function as the syntactic argument of a matrix predicate. But, in
the IHRC \ex{1}a, an internal head within the embedded clause
functions as its semantic argument whereas the embedded clausal
complement in \ex{1}b itself is the semantic argument of the
matrix predicate.


 \eenumsentence{\label{diff}
\item\shortex{5}
 { John-un & [Mary-ka & talli-nun & kes]-ul & {\bf capassta}. }
 { John-{\sc  top} & Mary-{\sc nom} & run-{\sc pne} & {\sc kes}-{\sc acc} & caught }
 {`John caught Mary who was running.'}
\item\shortex{5}
 { John-un & [Mary-ka & talli-nun & kes]-ul   & {\bf mollassta}. }
 { John-{\sc top} & Mary-{\sc nom} & run-{\sc pne} & {\sc kes}-{\sc acc} & not.know }
 {`John didn't know that Mary was running.'}}

%{kyongchal-i & [Tom-i   &somaycikiha-nun & kes]-ul & mollassta. }
% {police-{\sc nom}&  Tom-{\sc nom}& pickpocket-{\sc pne} & thing-{\sc acc} &  not.know }
% {`The police did not know Tom was pickpocketing.'}}

\noindent The only difference between \ex{0}a and \ex{0}b is the
matrix predicate. This difference induces the meaning difference.
As in \ex{0}a, when the matrix predicate is an action verb such as
{\it capta} `catch', {\it chepohata} `arrest', or {\it mekta}
`eat', we obtain entity readings. But as in \ex{0}b we have only
event readings when the matrix predicate is a type of recognition
verb such as {\it po-ta} `see', {\it al-ta} `know', and {\it
kiekhata} `remember'.

The key point in our analysis for the IHRC is that its
interpretation is dependent upon the type of matrix predicate.
What we assume is that the matrix predicate affects the
interpretation of the pronoun {\it kes}. In the lexical entry we
sketched in (\ref{entry2}), the subcategorization information of a
predicate involves not only syntax but also semantics. For
example, the verb {\it capassta} `caught'  in \ex{1} lexically
requires its object to refer to a referential individual whereas
the verb {\it mollassta} `not.know' in \ex{2} selects an object
complement whose index can refer to a propositional situation.


\enumsentence{
\begin{avm}
\[\q<capassta `caught'\q>\\
  ARG-ST \q<NP$_{i}$, NP$_{j}$\q>\\
  CONT\|RESTR \< \[RELATION &  catch\\
                            AGENT  &  {\it i}[MODE {\it ref}]\\
                   PATIENT & {\it j}[MODE {\it ref}]\]  \>\]\end{avm}}

\enumsentence{\begin{avm}
\[\q<mollassta `not.know'\q>\\
  ARG-ST \q<NP$_{i}$, NP$_{s }$\q>\\
  CONT\|RESTR \q< \[RELATION  & not.know\\
                         EXPERIENCER  &  {\it i}[MODE {\it ref}]\\
                         THEME  &  {\it s}[MODE {\it sit}]\]\q> \]
\end{avm}}

\noindent In the IHRC construction, such lexical requirements are
in one sense passed on to the head element {\it kes} and this
semantically empty pronoun will look for one (most) salient
discourse binder such as either an  individual as in \ex{-1}a or a
propositional event as in \ex{-1}b.

Such lexical requirements given in \ex{-1} above will ensure that
{\it kes} is properly coindexed with one of its semantic
restrictions. For example, our system would generate the structure
\ex{1}a for the sentence \ex{-2}a and the structure \ex{1}b for
the sentence \ex{-2}b:

\eenumsentence{
\item
\begin{tabular}[t]{cccccccccc}
                 \multicolumn{2}{c}{\node{1}{ VP}} \\
\\
                 \node{2}{\begin{avm}\avml \hfil NP$_{j}$\avmr\end{avm}}
                          &         & \node{3}{V}\\[1ex]
\\[1ex]

\node{4}{Mary$_{j}$-{\sc nom} run-{\sc pne} kes$_{j}$-ul} &
&\node{5}{caught}
\end{tabular}
\nodeconnect{1}{2}\nodeconnect{1}{3}\nodetriangle{2}{4}
\nodeconnect{3}{5}

\item
\begin{tabular}[t]{cccccccccc}
                 \multicolumn{2}{c}{\node{1}{ VP}} \\
\\
                 \node{2}{\begin{avm}\avml \hfil NP$_{s}$\avmr\end{avm}}
                          &         & \node{3}{V}\\[1ex]
\\[1ex]

\node{4}{Mary-{\sc nom} run$_{s}$-{\sc pne} kes$_{s}$-ul} &
&\node{5}{not.knew}
\end{tabular}
\nodeconnect{1}{2}\nodeconnect{1}{3}\nodetriangle{2}{4}
\nodeconnect{3}{5}
 }

%chan
%\noindent In the tree diagram in \ex{0}a, the RESTR value of {\it
%kes} is empty, since it is treated as an expletive pronoun, which
%has the empty restriction value (Sag and Wasow 1999). Otherwise,
%the NPj's RESTR value is <[1], [2], [1]>, which is wrong.
%chan

As represented in the structures, the meaning of {\it kes} in
\ex{0}a is identical with the noun phrase {\it Mary} whereas the
one in \ex{0}b is coindexed with the predicate {\it talli-nun}
`run-{\sc pne}'. This is possible due to the lexical entry for
{\it kes} given in (\ref{entry2}). Given this lexical entry, the
{\it kes} in \ex{0} will look like the following:

\enumsentence{
\begin{avm}\avml
\[\q<kes\q>\\
  HEAD \; {\it noun} \\
  ARG-ST \q<NP[{\it nom}]$_{j}$, V[VFROM {\it nun}]$_{s}$\q>\\
  CONTENT\|INDEX {\it j/s} \]
 \avmr

\end{avm}
}

\noindent As noted, the INDEX value of {\it kes} can be coindexed
with that of any element in its ARG-ST, either the subject
`Mary-{\sc nom}' or the predicate V `run-{\sc pne}'. This would
result in assigning a referential reading to the IHRC NP in
\ex{0}a as indicated by NP$_{j}$ whereas a situational (or event)
reading to the top NP in \ex{0}b as indicated by NP$_{s}$.

% The violation of
%this coindexation relation would violate the constraints in their
%lexical entries.

One clear advantage of such an analysis is a clean account of the
near complementary distribution of the clausal complement NP and
the IHRC, as well as for their structural identity, which no
analyses have paid attention to. The analysis obtains an entity
reading when the index value of {\it kes} identified with that of
an argument of the matrix predicate. Meanwhile, the analysis
induces an event reading for the IHRC when the index value is
structure-sharing with that of the adnominal predicate. This
analysis, thus, correctly predicts no cases where two readings are
available simultaneously.

The proposed analysis could also account for facts pertaining to
floating quantifiers. There is a kind of locality condition (e.g.,
mutual c-command relation) on the structural relationship between
a floating quantifier and its interpretively associated argument
NP:

\eenumsentence{\item \shortexnt{7}
 { Tom-un  &[{\bf sakwa-ka}& {\bf seykay-ka} & iss-nun & sangca-ul]}
 { Tom-{\sc top} & apple-{\sc nom}& three-{\sc nom} & exist-{\sc pne} & box-{\sc acc}}

\shortex{3} { hwumchessta.} { stole}{'Tom stole the box that had
three apples.'}

 \item \shortexnt{7}
 { *Tom-un  &[{\bf sakwa-ka} & iss-nun & sangca-ul] & {\bf seykay-ka}}
  { Tom-{\sc top} & apple-{\sc nom}& exist-{\sc pne} & box-{\sc acc} & three-{\sc
 nom}}

\shortexnt{2} {hwumchessta.} {stole}}

\noindent The data suggest that the host of a floating quantifier
can be only a nominal element in the same clause. Such a locality
condition in cases like \ex{0}b can be overridden in the IHRC as
in \ex{1}:

\enumsentence{ \shortexnt{5}
 { Tom-un  &[{\bf sakwa-ka}& cayngpanwi-ey & iss-nun & {\bf kes}]-ul}
 { Tom-{\sc top} & apple-{\sc nom}& tray-{\sc acc} & placed-{\sc pne} & {\sc kes}-{\sc acc}}

\shortex{2} {{\bf sey-kay-lul/*ka} & mek-ess-ta. }
 {three-{\sc cl}-{\sc acc}/{\sc nom} & eat-{\sc past}-{\sc decl}}
 { `Apples were on the tray, and Tom ate three of them.'}}

\noindent In the present analysis, the classifier {\it say kay}
`three things' in \ex{0} is construed with {\it kes} whose meaning
in turn is identical with that of the target {\it sakwa} `apple'.
This can be roughly represented as in \ex{1}:\footnote{Following
Sag and Wasow (1999), we assume that the mother's index value is
indentical with that of the head daughter.}

\enumsentence{{\small
\begin{tabular}[t]{cccccccccc}
  &         &  \multicolumn{2}{c}{\node{1}{ VP}} \\
             \\

            &   \multicolumn{2}{c}{\node{2}{\begin{avm}\avml \hfil NP$_{i}$\\
                           \avmr\end{avm}}}

        &    & \node{3}{V}\\
          \\
\multicolumn{2}{c}{\node{4}{\begin{avm}\avml \hfil NP$_{i}$\\
                           \avmr\end{avm}}}

&&\node{5}{\begin{avm}\avml \hfil N$'$$_{i}$\\
                           \avmr\end{avm}}  &\node{11}{ate}\\
          \\
 \node{6}{apple$_{i}$-{\sc nom} tray-on exist-{\sc pne}  kes-{\sc
 acc}}
 &&&\node{7}{say kay-{\sc acc}}

\end{tabular}
\nodeconnect{1}{2}\nodeconnect{1}{3}\nodeconnect{2}{4}
\nodeconnect{2}{5}\nodetriangle{4}{6}
\nodeconnect{5}{7}\nodeconnect{3}{11}}}


 \noindent
As can be seen from the structure, the pronoun {\it kes} is
identified with `apple' in the IHRC clause. The semantic index
value of this head pronoun NP is passed up to the NP construction.
The classifier is then construed with this NP whose index value is
again percolated up to the topmost NP which eventually is
coindexed with the PP in the clause.  One general constraint in
the language is that the case value of a floating quantifier
should match that of its host NP. Notice that the case marking on
the floating quantifier in \ex{0} cannot be nominative but must be
accusative. This once again supports our claim that the IHRC
construction is a nominal NP that can serve as the antecedent of a
floating quantifier. If not, we need to look for a different NP
with the accusative case in the same local domain.

One thing to notice here is that though we treat {\it kes} as a
kind of pronoun, the present analysis restricts its antecedent to
be within a restrict domain, neither deeply embedded within the
IHRC nor located outside the clause. This brings us one welcoming
result: it easily captures the fact that the IHRC construction
cannot be treated as an instance of the unbounded dependency as in
the EHRC construction:

\eenumsentence{
\item\shortexnt{7}
{Na-nun & [kutul-i &       [ \tr$_{i}$ unhayng-eyse &
nawassta-ko]}
 {I-{\sc top} &          they-{\sc nom} & bank-from
& came-out-{\sc comp}}

\shortex{4} {malha-n] & kangtoi-lul & capassta. (EHRC)} { say-{\sc
pne} & robber-{\sc acc} & caught} {`I arrested the robber who they
said was coming out of the bank.'}

\item\shortexnt{8}
{*Na-nun      & [kutul-i  &          [ kangtoi-ka  &
unhayng-eyse}{ I-{\sc top}  &           they-{\sc nom} &
robber-{\sc nom} & bank-from}

\shortex{5} {  nawassta-ko] &  malha-n] & kes$_{i}$-ul & capassta.
(IHRC)} { came-out-{\sc comp} & say-{\sc pne} & {\sc kes}-{\sc
acc} & caught}{`I arrested the robber who they said was coming out
of the bank.'}}

\noindent In the EHRC \ex{0}a, the head of the EHRC phrase, {\it
kangto}, is construed with the gap within the deeply embedded
clause as in the English relative clause. In \ex{0}b, however, the
head of the IHRC phrase, {\it kes}, cannot be construed with {\it
kangto}, showing that the IHRC phrase does not involve the
unbounded dependency.



\section{Pragmatic Aspects}
\subsection{Implicit Antecedent}

One of the remaining issues in the present analysis concerns cases
where the pronoun {\it kes} has an implicit antecedent whose data
we repeated here in \ex{1}:

\enumsentence{\shortex{5} { [[khep-uy & mwul-i & nemchi-n]  &
kes-ul]& ttakassta.}
 {cup-{\sc gen} &water-{\sc nom}&overflowed-{\sc pne}& {\sc kes}-{\sc acc}
 &wiped-out}
 {`(I) wiped out the water that overflowed from the one in the
 cup.'}}

\noindent In such cases, the antecedent of the pronoun {\it kes}
is an implicit participant resulted from the event denoted by the
IHRC clause.

Interestingly, such a phenomena can be observed in the so-called
pseudo relative clauses as in \ex{1} (See Kim 1998):

\eenumsentence{\item \shortex{3} {[mwul-i & hulu-nun]& soli}
{water-{\sc nom} &flow-{\sc pne}& sound} {`the sound of water's
flowing'}
\item\shortex{4}
{[komu-ka & tha-nun]& naymsay} {rubber-{\sc nom}   & burn-{\sc
pne} & smell} {`(literally) the smell such that rubber is
burning',} {`the smell that characterizes the burning of rubber'}}

\noindent What the sentence in \ex{0} describes is one of the
possibilities that could happen or result from the event of
water's flowing. Informally, such a meaning can be represented as
in \ex{1} (see Yoon 1993 also):

\enumsentence{
 $\lambda$$x$$[$sound$'$($x$) $\&$ flow$'$(w)
$\&$ perceptive-result-event(flow$'$(w), $x$)$]$}

\noindent There exist  the sound {\it x} and  the event of water's
flowing and this {\it x} is in the perceptive-event-relation with
the event of water's flowing.\footnote{Yoon's (1993) analysis
takes this `perceptive-result-event' relation as pragmatic
R-relation (relative clause relation) for all types of relative
clauses.  A support for such an interpretation could be found from
the fact that the appropriate paraphrasing of the psuedo relative
clause \ex{-1}b into the canonical relative clauses is something
like (i):

\enumsentence[(i)]{\shortex{4} {[[komu-ka & tha-lttay] & na-nun]&
naymsay} {rubber-{\sc nom}   & burn-when &come.out-{\sc pne} &
smell} {`the smell that comes out when rubber is burning'}}}

When there is no such perceptive-result relation between the
clause and the head, the pseudo relative clause is not acceptable:

\eenumsentence{
\item\shortex{4}
{[thayphwung-i &cinaka-n]   & huncek} {typhoon-{\sc nom}
&passed.by-{\sc pne}& debris} {`(literally) the debris such that a
typhoon passed by'} {`the debris that resulted from a typhoon's
passing by'}
\item\shortex{4}
{*[thayphwung-i &cinaka-n]   & phihay} {typhoon-{\sc nom}
&passed.by-{\sc pne}& damage} {`(intended) the damage caused from
a typhoon'}}

\noindent Though the debris could be a result of a typhoon we can
perceive, the abstract NP {\it phihay} `damage' is not.

We accept that such a pragmatic relation also holds in the IHRC
construction too. We may attribute such a pragmatic relation to
constructional constraints on the phrase that combines an
adnominal element with a limited set of head elements including
{\it kes}:\footnote{We assume that {\it head-adnom-comp-ph} has at
least two subtypes {\it head-pseudo-ph} and {\it head-ihrc-ph}
(cf.\ Kim 1998).}


\enumsentence{Constraints on {\it head-adnom-comp-ph}:\\

{\it head-adnom-comp-ph} \ \ $\rightarrow$ \ \

\hspace{0.5in}
\begin{avm}
    \[INDEX s\\
         ARG-ST \@a  $\oplus$ \<\[RELN {\it perceptive-result}\\
                                  ARG1  s\]\>

         \] \; \; {\bf H}[ \ \ ]

                            \end{avm}}
% chan

\noindent The constraint in \ex{0} tells us that in an instance of
{\it head-adnom-comp-ph}, the adnominal predicate (denoting a
situation {\it s}) can add to its ARG-ST an additional argument
denoting a salient participant (perceptive-result) {\it i}
generated from the situation which we obtain from the result of
the event {\it s}. This can be roughly represented in \ex{1}:


\enumsentence{

\begin{tabular}[t]{cccccccccc}
                 \multicolumn{2}{c}{\node{1}{N$_{i}$}} \\
\\
                 \node{2}{\begin{avm}\avml \hfil V\[INDEX s\\
                 ARG-ST \@a
                 $\oplus$ \q<[{\it per-result}]$_{i}$\q>\]
                 \avmr\end{avm}}
                          &         & \node{3}{N$_{i}$}\\[1ex]
\\[1ex]

\node{4}{...} & &\node{5}{kes}
\end{tabular}
\nodeconnect{1}{2}\nodeconnect{1}{3}\nodetriangle{2}{4}
\nodeconnect{3}{5}}

\noindent The notion of this context-based argument is similar to
a shadow argument in Pustejovsky (1998). Such an argument refers
to semantic content that is not necessarily expressed in syntax
and appears only by such pragmatic specifications (cf.\
Pustejovsky 1998).\footnote{Since such a constraint is specified
on the supertype of the pseudo relative clause and the IHRC, it
catpures the generalizations of the two types which otherwise we
would miss.}

The decision of the implicit argument is dependent upon various
grammatical factors: lexical, semantic, and pragmatic. For
example, when context prefers an overt element to be the
antecedent of {\it kes}, this explicit antecedent is preferred
over an implicit argument produced from the constraint in \ex{0}.
Consider the pairs in \ex{1}:

\eenumsentence{
\item\shortex{6}
{[paci$_{i}$-ka & telewe ci-n& kes$_{j}$-ul] & ttakanayssta}
{pants-{\sc nom}      & dirty & become & {\sc kes-acc} & wiped
out}{`The pants became dirty and (I) washed out the dirt from
them.'}
\item\shortex{6}
{[paci$_{i}$-ka & telewe ci-n& kes$_{i}$-ul] & ppalassta}
{pants-{\sc nom} & dirty & become & {\sc kes-acc} & washed}{`The
pants became dirty and I washed them.}}

\noindent Though the semantic argument of the matrix predicate in
\ex{0}a is an implicit argument, the one in \ex{0}b is the
subject. We cannot wipe out the pants themselves, but the dirt
itself  as in \ex{0}a. Though we can wash the dirt or the pants,
the preferred antecedent is an overt one, the subject.



\subsection{Relevancy Condition}

We accept the view that pragmatic conditions such as `relevance
condition or simultaneity condition' (Kuroda 1976) also play
important roles in selecting its own semantic restriction value.
For example, the relevancy condition specifies that an IHRC clause
should be interpreted as pragmatically in such a way as to be
directly relevant to the pragmatic content of its matrix clause
(see Uda 1998, Kim 2002 for detailed discussion). This condition
accounts for the following contrast:

\eenumsentence{
\item\shortexnt{9}
{Tom-un  &[sakwa-ka& cayngpan-uy-ey& iss-nun &kes]-ul}
 {Tom-{\sc top} & apple-{\sc nom}&tray-{\sc top}-{\sc loc} &exist-{\sc pne}
&{\sc kes}-{\sc acc}}

\shortex{2}{mekessta.}{ate} {`Tom ate the apple that was on the
tray.'}

\item\shortexnt{8}
{$\#$Tom-un &[sakwa-ka& eche   &  cayngpan-uy-ey &iss-ess-ten}
 {Tom-{\sc top} & apple-{\sc nom}& yesterday
&tray-{\sc top}-{\sc loc} & exist-{\sc pst-pne}}

\shortex{4}
 { kes]-ul & onul & mekessta. }
 { {\sc kes}-{\sc acc} & today &ate} {`Today Tom ate the apples, which were on the
tray yesterday.'}}

\noindent The difference between these two sentences is that in
the IHRC of \ex{0}b there is a time adverb {\it yesterday} and the
adnominal verb has the past form {\it iee-ess-ten}. The existing
condition is that the two events described by the matrix and the
embedded clause should be in the identical temporal location.

Observe that such a condition does not exist in the EHRC:

\enumsentence{\shortexnt{7} {Tom-un  &[ecey & cayngpan-uy-ey&
iss-ess-ten & sakwa]-ul}
 {Tom-{\sc top} & yesterday &
tray-{\sc top}-{\sc loc} &exist-{\sc pst}-{\sc pne} & apples-{\sc
acc}}

\shortex{3} {onul achim-ey  & mekessta.}
 { this morning &ate} {`This morning Tom ate the
apple that was on the tray yesterdy.'}}

\noindent In a similar fashion, As also claimed by Y.B. Kim
(2002), there appears to exist a strong `meaningful' relationship
between the IHRC and the matrix clause.\footnote{We leave open how
to formalize this condition within the HPSG feature system.
Informally, what we can say is, following Kim (2002), that the
event denoted by the IHRC is relevant to the matrix event when
both events are in the set of relations retrievable from the
background of the discourse participants. }

%chan 이 섹션 (relevancy condition)에서는 그냥 현상에 대한
%description만 있는것 같은데 차라리 빼면 안될까? 이런 relevancy
%condition이 왜 존재할까에 대한 대답이 있으면 좋겠는데... 이 문제는
%좀더 생각해 봐야겠다. focus랑 뭔가가 연결 될 것도 같은데 이건 좀
%더 생각을 해 봐야겠다.%chan


\section{Information Packaging and More on the Differences}



One telling property that differentiates the IHRC from the EHRC is
that unlike the EHRC, the IHRC cannot function as an answer to a
{\it wh}-question, as we observe in \ex{1}:

\enumsentence{\label{wh-question} \shortex{5} {A:  kyongchal-i  &
nwukwu-lul & capasstako?} {\hspace{0.25in}  policeman-{\sc nom}&
who-{\sc acc} & caught} {\hspace{0.25in}`Who did you say the
policeman caught?'}

\shortex{6} {B: [[unhayng-eyse & nao-nun] & kangto-lul] &
capasse.} {\hspace{0.25in} bank-from & come.out-{\sc pne} &
robber-{\sc acc} & caught} {\hspace{0.25in}`(They) caught the
robber coming out from the bank.'}

\shortexnt{6} {B$'$: $\#$kangto-ka & unhangy-eyse & nao-nun &
kes-ul & capasse. } {\hspace{0.25in} robber-{\sc nom}& bank-from &
come.out-{\sc pne} & {\sc kes}-{\sc acc} & caught}}

%\noindent This appears to be related with the fact that the IHRC
%construction cannot be used in isolation as a simple referring
%expression:\footnote{This differentiates IHRCs from sentential
%nominalizations.}
%
%\enumsentence{ \shortex{4} {*kangto-ka & unhangy-eyse & nao-nun &
%kes.} {robber-{\sc nom}& bank-from & come.out-{\sc pne} & {\sc
%kes}} {`the thing that a robber is coming out from the bank.'} }

%\noindent What this tells us is that the IHRC construction does
%not serve as an expression
% referring to an individual by itself.

This paper claims that such a difference between the IHRC and the
EHRC are basically due to what is focused: In the IHRC, the event
described by the IHRC clause, that is, denoted by the adnominal
predicate, is newly conveyed information whereas in the EHRC no
such a restriction holds. We could attribute this as a lexical
constraint on the pronoun {\it kes} as represented in \ex{1} (cf.\
Engdahl and Vallduv\'{\i} (1996)):

%chan IHRC clause가 newly conveyed information이 아니라는 생각도 드네. 예를 들어,
%
%(TV 모니터를 들여다 보며)
%A: 어 도둑이 벌써 은행에서 나오네.
%B: 걱정마. 경찰이 [___ 나오는 것을] 잡을 거야.%
%
%이경우, IHRC는 이미 new information은 아니지 않나? 만일 이렇다면 이건 큰 문젠데...원래 생각과 너무 달라지게 되지 않나? 어찌 생각하냐? 내가 아직 focus 쪽으로 감이 잘 잡히지 않네. %chan
%
%the focused element is not the
%event itself that the relative clause describes but the individual
%referred to by the relative head.


\enumsentence{Lexical Entry for {\it kes} (final):\\
\begin{avm}\avml
\[\q<kes\q>\\
  HEAD \; {\it noun} \\
  ARG-ST \@b\q<...[\;]$_{i}$,... [\;]$_{s}$...\q>\\
  CONTENT\|INDEX {\it i/s}\\
  INFO-ST\|FOCUS {\it s}\]\\

{\footnotesize (where \@b results from the argument composition
\<V\[FORM {\it (n)un}\\
  ARG-ST \@a\]\> $\oplus$ \@a)}
\avmr

\end{avm}
}

\noindent The lexical entry means that the pronoun {\it kes}
constructionally assigns focus value to the preceding adnominal
predicate. What this implies is that the embedded IHRC clause
carries focus and conveys new information about the event
represented by the clause predicate. That is, following Ohara
(1996), we also accept the view that the IHRC clause has a
function of event reporting.) We could observe that an IHRC cannot
be used when the event described by the IHRC is given in a
previous context. For example, when the information such that John
was eating apples was already given in prior context, unlike the
EHRC in \ex{1}b, the IHRC in \ex{1}c is an inappropriate
statement:

%chan Align the glosses and interpretation.
%Add one more sentence at the end of the (a) sentence to make the judgment clearer:

%Ku   sakwa-nun  apeci-uy     kwaswuwon-eyse   tta-n       kes-iessta.
%The  apple-Top   father-Gen  orchard-at  pick-Adm   thing-past.copula
%'The apple was picked at his father's orchard.'
%chan

\eenumsentence{
\item\shortex{10}
{[A:] &.... John-i & sakwa-lul& mek-ko & issessta...} {&John-{\sc
nom} &apple-{\sc acc} & eat-{\sc comp} & in.progessive}{`...John
was eating an apple...'}

\item\shortexnt{10}
{[B:] & kuttay & kapcaki &nwukwunka-ka &[sakwa-lul& mek-ko}
 { & then& suddenly &someone-{\sc nom} &apple-{\sc acc}&
eat-{\sc comp}}

\shortex{3} {iss-nun] & John-ul &pwulessta} { is-{\sc pne} &
John-{\sc acc}& called}{`Then suddenly somebody called John, who
was eating apples.'}

\item\shortexnt{10}
{[B:]& $\#$kuttay &kapcaki &nwukwunka-ka & [John-i &sakwa-lul}
  {&then& suddenly &someone-{\sc nom} & John-{\sc
nom} &apple-{\sc acc}}

\shortexnt{5}{ mek-ko& iss-nun] & kes-ul& pwulessta} {eat-{\sc
comp} &is-{\sc pne} & {\sc kes-acc}& called}}


\noindent In our analysis, \ex{0}b is not a natural continuation
when the information that John's eating the apples is given
information. This is because the event of John's eating an apple
has already been introduced and cannot function as carrying new
information. In this sense, the IHRC represents `information
focus', conveying new, nonpresupposed information without
expressing exhausitive identification performed on a set of
contextually or situationally given entities (cf. Kiss 1998).

This implies that the IHRC construction cannot serve as an
expression referring to an individual, but can function  only as a
reply to an event asking query. Such a fact can be attested by
another example. An IHRC can be an answer only to an event asking
query like \ex{1}A:

\eenumsentence{\item[A:]\shortex{5} {kyongchal-i & totuk-ul &
etteskey &capasstay?} {police-{\sc nom} & robber& how & caught?}
{(Do you know) how the police caught the robber?'}

\item[B:]\shortexnt{7}
 {Kyongchal-i & [totwuk-i & ton-ul & hwumchi-nun &  kes-ul]}
 {police-{\sc nom} &  thief-{\sc nom}& money  &    steal-{\sc pn}  &  {\sc kes}-{\sc acc}}

\shortex{3}
 {capass-tay. }
{ caught-said}
 {`(People) said that the police arrested the robber who was stealing money.'}
 }


\noindent Once we accept the proposed view, we can provide a
streamlined analysis for several complicated properties of the
IHRC construction as well as the differences between IHRCs and
EHRCs. As we have seen earlier, various phenomena indicate that
the IHRC construction has some nominal properties: nominal case
markings and passivization. However, unexpected from these nominal
properties, we cannot cleft the construction as in \ex{1} because
of the mismatch in what is focused.

\eenumsentence{\label{cleft} \item\shortex{6} {[Mary-ka & cap-un &
kes]-un & talli-nun & [Tom-i-ta].} {Mary-{\sc nom}& catch-{\sc
pne} &{\sc kes}-{\sc pne}& run-{\sc pne}& Tom-{\sc nom} {\sc
cop}-{\sc decl}} {`(int.) What Mary caught was Tom, who was
running.'}

\item\shortexnt{6}{*[Mary-ka & cap-un  &
kes]-un & [Tom-i & talli-nun} {\hspace{0.05in} Mary-{\sc nom}&
catch-{\sc pne} &{\sc kes}-{\sc pne}& Tom-{\sc nom}&run-{\sc pne}}

\shortex{3} {kes]-i-ta.} {{\sc kes}-{\sc cop}-{\sc decl}} {`(int.)
What Mary caught was Tom, who was running.'}}

\noindent As a canonical constraint on the cleft-construction, the
focused value cannot be a VP or an event. The canonical focused
value is an NP nominal. Under our assumption, the IHRC
construction, though syntactically an NP, focuses an event whereas
the focused element in the cleft is generally an NP referring to
an individual.



Another welcoming consequence of the analysis is that it can
provide a clue as to why it is not possible to have an
unaccusative verb or a verb in IHRC that describes an intrinsic
property of an entity as in \ex{1}: The most natural class of verb
that can report an event or describe an event is a stage level
predicate.\footnote{The EHRC counterpart is grammatical.}

\enumsentence{ \shortexnt{7}
 { *Tom-un  &[John-i & {\bf hyonmeyongha-n} & kes]-ul}
  {  Tom-{\sc top} & John-{\sc nom}& smart-do-{\sc pne}& {\sc kes}]-{\sc acc}}

\shortex{3} {chochenghayessta. }
     {invited }
 { `(int.) Tom invited John, who was smart.'}}


It has been also noted that the IHRC cannot be in the form of
negative as in \ex{1}a.

\eenumsentence{
\item\shortex{5}
 {$\#$John-i & [[Tom-i       & {\bf an talli-nun}]    &  kes]-ul & capassta.}
 {John-{\sc nom} &Tom-{\sc nom} & not run-{\sc pne}   &   {\sc kes}-{\sc acc} & caught}
 {`John caught Tom, who was not running.'}

 \item\shortexnt{6}
{John-i & [[Tom-i & {\bf memcwuci}  & {\bf anh-nun}]    &
kes]-ul}
 {John-{\sc nom} &Tom-{\sc nom} &stop  &  not-{\sc pne}   &  {\sc kes}-{\sc acc}}

\shortex{3} {capassta.}
  {caught}
 {`John caught Tom, who wasn't stopping.'}}


Such a condition can also be found in English locative inversion:

\enumsentence{a. *On the wall never hung a picture of U.S.
Grant.\\
b. On the wall hangs not a picture of U.S. Grant but one of
Jefferson Davis. (Aissen 1975)}

\noindent According to Aissen (1975), the locative phrase
functions as a backdrop, and the assertion that such a scene does
not exist cannot serve this purpose. We conjecture that such a
condition also holds in the Korean IHRC.  Within our theory, this
is so because there is no event to be focused. But if the negative
IHRC entails an event that is happening or happened, we could
focus the IHRC construction as shown in \ex{0}b. The IHRC in
\ex{0}b entails that Tom remained as he was and John caught him.
Thus what is focused would be the semantic content of the IHRC
that includes a resultant event from the clause.


In sum, what the present analysis shows us is that the IHRC is
syntactically and semantically a nominal construction whereas in
terms of information packaging it has sentence-like properties in
that the event described by the clause is focused.



\section{Conclusion}

We have shown that the Korean IHRC is formed by a
complex-predicate mechanism of the semantically empty pronoun,
{\it kes}: the pronoun combines with an adnominal verb, forming a
strong syntactic unit. And the selection of the internal head is
dependent upon the semantics of the matrix predicate and context
in question.

We have also claimed that the IHRC reading is obtained when the
pronominal {\it kes} is coindexed with one argument of the
adnominal verb. Meanwhile, we obtain an event reading when the
pronominal is coindexed with the eventive relation of the
adnominal verb. The present analysis claiming what is focused
differentiates between the IHRC and the EHRC provides a clean
account of their differences in various phenomena. This line of
lexicalist, nonderivational analysis could avoid the postulation
of any phantom formatives (such as {\it pro}), and eventuallly
provides us with a clearer and simpler grammar of Korean (and
possibly Japanese too).

\section*{Acknowledgements}

An earlier version of this paper was presented at the 9th
International Conference on HPSG held at Kyung Hee University,
Seoul from August 6 to 7. We are grateful to Andreas Kathol,
Sae-Youn Cho, Chungmin Lee, Byung-Soo Park, Byung-Rae Ryu, and
Peter Sells  for criticisms and comments. All the remaining errors
are of course ours. This work was supported by the Korea Research
Foundation (grant number 2001-042-A00037).



{
\section*{References}
}
\begingroup
 \parindent=-2em \advance\leftskip by2em
 \parskip=0.5ex
\par\indent\vspace*{-15pt}

Aissen, Judith. 1975. Presentational-{\it there} insertion: A
cyclic root transformation. {\it Chicao Linguistic Society} 11:
1--14.



Bratt, Elizabeth. 1995. {\it Argument Composition and the Lexicon:
Lexical and Periphrastic Causatives in Korean}. Doctoral
Dissertation. Stanford University.

%chan If we adopt the constraint on the index and the flat-clause structure, add the follwoing:

Chung, Chan. 1998. Argument composition and long-distance scrambling in Korean: an extension
of the complex predicate analysis.
 {\it Syntax and Semantics} 30: 159-220. Academic Press: New York.

%chan

Chung, Chan. 1999. Complex noun, Multiple inheritance and
Internally Headed Relativization in Korean. {\it BLS} 29.

Chung, Dae-Ho. 1996. On the structure of the so-called head
internal relative Construction. {\it Proceedings of the Pacific
Asia conference on Language, Information and Computation}, 11:
393--402.

Cole, Peter. 1987. The Structure of Internally Headed Relative
    Clauses. {\it Natural Language and Linguistic Theory} 5, 277-302.

Culy, Christopher.   1990. {\it The Syntax and Semantics of
IHRCs.} Doctoral Dissertation. Stanford University.

Engdahl, Elizabeth and Enric Vallduv\'{\i}. 1996. Information
packaging in HPSG. In Claire Grover and Enric Vallduvi (eds.),
{\it Studies in HPSG}: Edinburgh Working Papers in Cognitive
Science, 1--31.

Hoshi, Koji. 1996. The head-internal relative clause in Japanese:
An empty head noun approach. {\it Japanese and Korean
Linguistics}, 5: 253--268.

Ito, Junko. 1986.  Head-Movement at LF and PF: The Syntax of
   Head-Internal Relatives in Japanese.  In {\it University of Massachusetts
   Occasional Papers in Linguistics 11}.


Jhang, Sea-Eun. 1991. Internally-headed Relative Clauses in
Korean. {\it Harvard Studies in Korean Linguistics}, 4: 269-280.

Kathol, Andreas and Ken Vanbik. 1999. Morphology-Syntax Interface
in Lai Relative Clauses. In Pius Tamanji, Masako Hirotami, and
Nancy Hall (eds.) {\it Proceedings of the 29th Annual Meeting of
the North Eastern Linguistic Society}, GLSA, University of
Massachusetts, Amherst. pp. 427--441.


Kim, Yong-Beom. 2002. Relevance in internally headed relative
clauses in Korean. {\it Lingua} 112: 541--559.

Kim, Jong-Bok. 1999. Grammatical Interfaces in Korean Internally
Headed Relative Clause Construction. {\it Language Research} 17:
257--275.

Kim, Jong-Bok. 1998. On the Mixed Properties of Pseudo Relative
Clauses. In B.S. Park and J. Yoon (eds), {\it
          Proceedings of the 11th International Conference on Korean Linguistics}, 83-93. ICKL

Kiss, Katalin. 1998. Identificational Focus versus Information
Focus. {\it Language} 74: 245--273.

Kuroda, S.-Y. 1976. Headless Relative Clauses in Modern Japanese
and the Relevancy Condition. {\it BLS} 7.

Murasugi, Keiko. 1994. Head-internal Relative Clauses as Adjunct
Pure Complex NPs. In Shuji Chiba (ed.), {\it Synchronic and
Diachronic Approaches to Langauge}, 425--437.

Ohara, K. H. 1996. {\it A Constructional Approach to Japanese
Internally-Headed Relativization.}. Ph.D. Dissertation. UC
Berkeley.

Park, Byung-Soo. 1994. Modification vs. Complementation: The
so-called internally headed relative clauses reconsidered. In
Ishikawa, A and Y. Nitta (eds.), {\it Proceedings of the 1994
Kyoto Conference: A Festschrift for Professor Akira Ikeya.} The
Logico-Linguistic Society of Japan, Kyoto, Japan.

Pustejovsky, James. 1998. {\it The Generative Lexicon.} Cambridge,
Massachusetts: The MIT Press.

Sag, Ivan and Tom Wasow. 1999. {\it Syntactic Theory: A Formal
Approach}. Stanford: CSLI Publications.


Sells, Peter. 1995. Korean and Japanese morphology from a lexical
perspective. {\it Linguistic Inquiry} 26.2.     277-325.


Uda, Chihara. 1998. A Multiple Inheritance Analysis of the
Internally-headed Relative Clause in Japanese. {\it Proceedings of
the Pacific Asia Conference on Language, Information and
Computation}, 12: 82--93

Watanabe, Akira. 1992. Subjacency and s-structure movement of
wh-in-situ. {\it Journal of East Asian Linguistics} 1: 255-291.

Yoon, Jae-Hak. 1993. Different Semantics for Different syntax:
Relative Clauses in Korean. {\it OSUWPL} 42: 199-226.

\endgroup
\end{document}